---
title: "01 — Data cleaning"
author: "Alexander Hobbs"
output: html_document
---

<!--
This script applies the data exclusion criteria described in the dissertation
methods to produce the cleaned analysis-ready dataset used by all
downstream scripts.

Inputs (in data/):
  - Blue_tits_raw.csv     : raw nest-level Phenoweb data
  - site_details.csv      : site coordinates and metadata

Outputs (written to data/):
  - Blue_tits_cleaned.csv : after duplicate removal and clutch-swap exclusion
  - Blue_tits_final.csv   : after additional analysis-ready filtering
                            (clutch size <= 16, etc.)
  - FED_Blue_tits_exc_2020.csv : same as Blue_tits_final but with 2020
                                 excluded (used for lay-date analyses)
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r libraries}
library(tidyverse)
library(geosphere)
```

# 1. Load raw data

```{r load-raw}
# Adjust filename to match whatever the raw export from the Phenoweb is called.
Blue_tits <- read.csv("data/Blue_tits_raw.csv",
                      stringsAsFactors = TRUE, header = TRUE)

cat("Raw data:", nrow(Blue_tits), "rows\n")
```

# 2. Remove duplicate nest records

Where the same nest box appears more than once in the same year, retain the
record with the earliest first-egg date (`fed`).

```{r remove-duplicates}
Blue_tits_cleaned <- Blue_tits %>%
  arrange(year, site, box, fed) %>%
  distinct(year, site, box, .keep_all = TRUE)

# Check no duplicates remain
n_remaining_dupes <- Blue_tits_cleaned %>%
  group_by(year, site, box) %>%
  filter(n() > 1) %>%
  nrow()
cat("Duplicates remaining:", n_remaining_dupes, "(should be 0)\n")
```

# 3. Remove clutch-swap experimental nests

Nests with `ClutchSwapTreatment == 1` were part of an experimental manipulation
(2017–2019) and have their fledgling success values set to `NA` so they
do not contribute to natural-variation estimates.

```{r exclude-clutch-swap}
Blue_tits_cleaned <- Blue_tits_cleaned %>%
  mutate(suc = ifelse(ClutchSwapTreatment == 1, NA, suc))
```

# 4. Apply analysis-ready filters

Exclude the Tomatin (TOM) site (insufficient occupancy across years) and
clutches > 16 eggs (biologically implausible — typically two-female clutches).

```{r analysis-filters}
Blue_tits_final <- Blue_tits_cleaned %>%
  filter(site != "TOM",
         is.na(cs) | cs <= 16)

cat("Final dataset:", nrow(Blue_tits_final), "rows across",
    n_distinct(Blue_tits_final$site), "sites and",
    n_distinct(Blue_tits_final$year), "years\n")
```

# 5. Build lay-date subset excluding 2020

Lay-date analyses exclude 2020 due to reduced fieldwork during COVID-19,
which produced a non-representative subset of late nests.

```{r exclude-2020-fed}
FED_Blue_tits_exc_2020 <- Blue_tits_final %>%
  filter(year != 2020)

# Add site-year combination column used in downstream scripts
FED_Blue_tits_exc_2020$siteyear <- paste(FED_Blue_tits_exc_2020$site,
                                          FED_Blue_tits_exc_2020$year)
```

# 6. Per-trait sample sizes

```{r sample-sizes}
Blue_tits_final %>%
  summarise(
    n_lay_date          = sum(!is.na(fed)),
    n_clutch_size       = sum(!is.na(cs)),
    n_fledgling_success = sum(!is.na(suc))
  )
```

# 7. Save cleaned datasets for downstream scripts

```{r save-outputs}
write.csv(Blue_tits_cleaned,        "data/Blue_tits_cleaned.csv",        row.names = FALSE)
write.csv(Blue_tits_final,          "data/Blue_tits_final.csv",          row.names = FALSE)
write.csv(FED_Blue_tits_exc_2020,   "data/FED_Blue_tits_exc_2020.csv",   row.names = FALSE)

cat("Saved 3 files to data/\n")
```
