---
title: "02 — Data summary statistics"
author: "Alexander Hobbs"
output: html_document
---

<!--
Reports overall and trait-specific sample sizes, completeness statistics,
and the data-driven justification for excluding the Tomatin (TOM) site
and the 2020 lay-date data. All summaries reported in the dissertation
methods (§2.2) and results (§3) sections.

Inputs (from data/, produced by 01_data_cleaning.Rmd):
  - Blue_tits_final.csv
  - Blue_tits_cleaned.csv
  - FED_Blue_tits_exc_2020.csv
  - Pairwise_distance_km.csv (produced by 03_pairwise_distances.Rmd)

For per-site-year aggregates (site_means, FED_site_means_exc2020),
this script expects them to exist in the workspace from a prior
session, OR can be regenerated from the inputs above.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r libraries}
library(tidyverse)
library(knitr)
```

```{r load-data}
Blue_tits_final        <- read.csv("data/Blue_tits_final.csv",
                                   stringsAsFactors = FALSE)
Blue_tits_cleaned      <- read.csv("data/Blue_tits_cleaned.csv",
                                   stringsAsFactors = FALSE)
FED_Blue_tits_exc_2020 <- read.csv("data/FED_Blue_tits_exc_2020.csv",
                                   stringsAsFactors = FALSE)
Pairwise_distance_km   <- read.csv("data/Pairwise_distance_km.csv",
                                   stringsAsFactors = FALSE)

# Build site-level aggregates needed in summary sections below
site_means <- Blue_tits_final %>%
  group_by(site, year) %>%
  summarise(mean_cs  = mean(cs,  na.rm = TRUE),
            mean_suc = mean(suc, na.rm = TRUE),
            .groups  = "drop")

FED_site_means_exc2020 <- FED_Blue_tits_exc_2020 %>%
  group_by(site, year) %>%
  summarise(median_fed = median(fed, na.rm = TRUE),
            .groups    = "drop")

FED_site_medians_exc2020 <- FED_site_means_exc2020
```

## 1. Overall sample sizes

```{r sample-sizes}
n_nests   <- nrow(Blue_tits_final)
n_sites   <- n_distinct(Blue_tits_final$site)
n_years   <- n_distinct(Blue_tits_final$year)
n_fed     <- sum(!is.na(FED_Blue_tits_exc_2020$fed))
n_cs      <- sum(!is.na(Blue_tits_final$cs))
n_suc     <- sum(!is.na(Blue_tits_final$suc))

cat("Total nests:", n_nests, "\n",
    "Sites:", n_sites, "| Years:", n_years, "\n",
    "FED obs:", n_fed, "| CS obs:", n_cs, "| SUC obs:", n_suc, "\n")
```

## 2. Per-site-year sample sizes

```{r site-year-sizes}
site_year_n <- Blue_tits_final %>%
  group_by(site, year) %>%
  summarise(n = n(), .groups = "drop")

cat("Nests per site-year — min:", min(site_year_n$n),
    "| max:", max(site_year_n$n),
    "| mean:", round(mean(site_year_n$n), 1), "\n")
```

## 3. Annual coefficient of variation per trait

```{r annual-cv}
annual_cv <- bind_rows(
  FED_site_medians_exc2020 %>%
    group_by(year) %>%
    summarise(cv = sd(median_fed, na.rm = TRUE) / mean(median_fed, na.rm = TRUE),
              .groups = "drop") %>%
    mutate(trait = "FED"),
  site_means %>%
    group_by(year) %>%
    summarise(cv = sd(mean_cs, na.rm = TRUE) / mean(mean_cs, na.rm = TRUE),
              .groups = "drop") %>%
    mutate(trait = "CS"),
  site_means %>%
    group_by(year) %>%
    summarise(cv = sd(mean_suc, na.rm = TRUE) / mean(mean_suc, na.rm = TRUE),
              .groups = "drop") %>%
    mutate(trait = "SUC")
)

annual_cv %>%
  group_by(trait) %>%
  summarise(mean_cv = round(mean(cv), 3), .groups = "drop") %>%
  kable(caption = "Mean coefficient of variation across years, per trait")
```

## 4. Trait summary table

```{r trait-summary}
trait_summary <- tibble(
  Trait   = c("First egg date", "Clutch size", "Fledgling success"),
  n       = c(n_fed, n_cs, n_suc),
  Mean    = c(mean(FED_Blue_tits_exc_2020$fed, na.rm = TRUE),
              mean(Blue_tits_final$cs,  na.rm = TRUE),
              mean(Blue_tits_final$suc, na.rm = TRUE)),
  Median  = c(median(FED_Blue_tits_exc_2020$fed, na.rm = TRUE),
              median(Blue_tits_final$cs,  na.rm = TRUE),
              median(Blue_tits_final$suc, na.rm = TRUE)),
  SD      = c(sd(FED_Blue_tits_exc_2020$fed, na.rm = TRUE),
              sd(Blue_tits_final$cs,  na.rm = TRUE),
              sd(Blue_tits_final$suc, na.rm = TRUE)),
  Min     = c(min(FED_Blue_tits_exc_2020$fed, na.rm = TRUE),
              min(Blue_tits_final$cs,  na.rm = TRUE),
              min(Blue_tits_final$suc, na.rm = TRUE)),
  Max     = c(max(FED_Blue_tits_exc_2020$fed, na.rm = TRUE),
              max(Blue_tits_final$cs,  na.rm = TRUE),
              max(Blue_tits_final$suc, na.rm = TRUE))
) %>%
  mutate(across(Mean:Max, ~ round(.x, 2)))

kable(trait_summary,
      caption = "Summary statistics for reproductive traits across all site-years.")
```

## 5. Nest failure rate

```{r nest-failures}
nest_failure_summary <- Blue_tits_final %>%
  drop_na(suc) %>%
  summarise(
    total_nests      = n(),
    failed_nests     = sum(suc == 0),
    failure_rate_pct = round((failed_nests / total_nests) * 100, 2)
  )

print(nest_failure_summary)
```

## 6. Distance summary

```{r distance-summary}
cat("Site pairs:", nrow(Pairwise_distance_km), "\n",
    "Distance range:", round(min(Pairwise_distance_km$distance_km), 1),
    "-", round(max(Pairwise_distance_km$distance_km), 1), "km\n",
    "Mean:", round(mean(Pairwise_distance_km$distance_km), 1), "km\n",
    "Median:", round(median(Pairwise_distance_km$distance_km), 1), "km\n")
```

## 7. Site-year coverage by trait and year

```{r coverage}
coverage <- Blue_tits_final %>%
  group_by(year) %>%
  summarise(
    sites_with_fed = n_distinct(site[!is.na(fed)]),
    sites_with_cs  = n_distinct(site[!is.na(cs)]),
    sites_with_suc = n_distinct(site[!is.na(suc)]),
    .groups        = "drop"
  )

kable(coverage,
      caption = "Number of sites with data per year for each trait.")
```

## 8. Year range and site-year completeness

```{r completeness}
cat("Year range:", min(Blue_tits_final$year), "-",
    max(Blue_tits_final$year), "\n")

total_possible   <- n_sites * n_years
actual_siteyears <- n_distinct(paste(Blue_tits_final$site,
                                     Blue_tits_final$year))
cat("Possible site-years:", total_possible, "\n",
    "Actual site-years:", actual_siteyears, "\n",
    "Completeness:", round(100 * actual_siteyears / total_possible, 1), "%\n")
```

## 9. Years per site (justifies TOM exclusion)

```{r tom-justification}
years_per_site <- Blue_tits_cleaned %>%
  filter(cs <= 14 | is.na(cs)) %>%
  group_by(site) %>%
  summarise(n_years = n_distinct(year), .groups = "drop")

cat("Years per site — min:", min(years_per_site$n_years),
    "| max:", max(years_per_site$n_years),
    "| mean:", round(mean(years_per_site$n_years), 2), "\n")

years_per_site %>%
  arrange(n_years) %>%
  head(5) %>%
  kable(caption = "Sites with fewest years of data")
```

## 10. Post-exclusion sample sizes

```{r post-exclusion}
cat("\n--- After excluding TOM ---\n")
cat("Sites:", n_distinct(Blue_tits_final$site), "\n")
cat("Nests:", nrow(Blue_tits_final), "\n")

cat("\n--- FED after excluding TOM + 2020 ---\n")
cat("FED obs:", sum(!is.na(FED_Blue_tits_exc_2020$fed)), "\n")
cat("Sites with FED:", n_distinct(FED_Blue_tits_exc_2020$site[!is.na(FED_Blue_tits_exc_2020$fed)]), "\n")
cat("Years:", n_distinct(FED_Blue_tits_exc_2020$year[!is.na(FED_Blue_tits_exc_2020$fed)]), "\n")
```

## 11. Per-trait site-year counts entering each analysis

```{r per-trait-counts}
siteyear_per_trait <- tibble(
  Trait = c("FED (excl. 2020)", "CS", "SUC"),
  `Site-year combinations` = c(
    FED_site_means_exc2020 %>% drop_na(median_fed) %>% nrow(),
    site_means %>% drop_na(mean_cs)  %>% nrow(),
    site_means %>% drop_na(mean_suc) %>% nrow()
  ),
  `Sites` = c(
    FED_site_means_exc2020 %>% drop_na(median_fed) %>% pull(site) %>% n_distinct(),
    site_means %>% drop_na(mean_cs)  %>% pull(site) %>% n_distinct(),
    site_means %>% drop_na(mean_suc) %>% pull(site) %>% n_distinct()
  ),
  `Years` = c(
    FED_site_means_exc2020 %>% drop_na(median_fed) %>% pull(year) %>% n_distinct(),
    site_means %>% drop_na(mean_cs)  %>% pull(year) %>% n_distinct(),
    site_means %>% drop_na(mean_suc) %>% pull(year) %>% n_distinct()
  )
)

kable(siteyear_per_trait,
      caption = "Site-year combinations entering each synchrony analysis (after all exclusions).")
```

## 12. 2020 lay-date breakdown (supports exclusion decision)

```{r 2020-fed}
cat("\n--- 2020 FED breakdown ---\n")
Blue_tits_final %>%
  filter(year == 2020) %>%
  summarise(
    total_nests_2020  = n(),
    sites_any_data    = n_distinct(site),
    fed_obs           = sum(!is.na(fed)),
    sites_with_fed    = n_distinct(site[!is.na(fed)]),
    sites_without_fed = n_distinct(site) - n_distinct(site[!is.na(fed)])
  ) %>%
  kable(caption = "2020 data summary: justification for FED exclusion")
```
