---
title: "05 — Synchrony estimates (ICCs)"
author: "Alexander Hobbs"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: flatly
---

<!--
Fits linear mixed-effects models to estimate intraclass correlation
coefficients (ICCs) for lay date, clutch size, and fledgling success.
Reproduces Tables 1, 2 and Figures 2, 3 in the dissertation.

Sequence:
  1. Baseline LMMs with year, site, site:year random effects -> Table 1
  2. Cascade models with upstream traits as fixed effects   -> Table 2 / Figure 3
  3. Temporal trajectory plots                              -> Figure 2

Inputs (data/, produced by 01_data_cleaning.Rmd and earlier scripts):
  - Blue_tits_final.csv
  - Blue_tits_cleaned.csv
  - FED_Blue_tits_exc_2020.csv
  - site_means.csv, site_details.csv, climate_data.csv
  - Pairwise_correlations.csv (precomputed pairwise Pearson r between sites)
  - site_means_with_temp.csv

Bootstrapping: parametric bootstraps (B = 5000) cached as
  icc_boot_results.rds and t2_boot_results.rds — delete to regenerate.
  Random seed 42 set at the top of each bootstrap chunk for reproducibility.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      fig.align = "center", dpi = 300)
```

```{r load-libraries}
library(tidyverse)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(broom)
library(knitr)
library(kableExtra)
library(corrr)
library(reshape2)
library(ncf)
library(mvtnorm)
library(matrixStats)
library(progress)
library(patchwork)
library(scales)
library(ggforce)
library(viridis)
library(geosphere)
library(lmeresampler)
```

***

# Data loading and cleaning

```{r load-data}
site_means <- read.csv(
  "data/site_means.csv",
  stringsAsFactors = FALSE, header = TRUE)

site_details <- read.csv(
  "data/site_details.csv",
  stringsAsFactors = FALSE, header = TRUE)

Blue_tits_cleaned <- read.csv(
  'data/Blue_tits_cleaned.csv',
  stringsAsFactors = TRUE, header = TRUE)

Blue_tits_final <- read.csv(
  'data/Blue_tits_final.csv',
  stringsAsFactors = TRUE, header = TRUE)

FED_Blue_tits_exc_2020 <- read.csv(
  'data/FED_Blue_tits_exc_2020.csv',
  stringsAsFactors = TRUE, header = TRUE)

Pairwise_correlations <- read.csv(
  "data/Pairwise_correlations.csv",
  stringsAsFactors = FALSE, header = TRUE)

site_means_with_temp <- read.csv(
  "data/site_means_with_temp.csv",
  stringsAsFactors = FALSE, header = TRUE)

climate_data <- read.csv(
  "data/climate_data.csv",
  stringsAsFactors = FALSE, header = TRUE)
```

```{r data-cleaning}
site_means <- site_means %>% filter(site != "TOM")
site_details <- site_details %>% filter(site != "TOM")

FED_site_medians_exc2020 <- site_means %>% filter(year != 2020)
FED_Blue_tits_exc_2020 <- Blue_tits_final %>% filter(year != 2020)
FED_Blue_tits_exc_2020$siteyear <- paste(FED_Blue_tits_exc_2020$site, 
                                          FED_Blue_tits_exc_2020$year)
site_details_clean <- site_details %>%
  dplyr::select(site, Mean.Lat, Mean.Long) %>%
  dplyr::rename(Lat = Mean.Lat, Lon = Mean.Long)

site_temp_summary <- climate_data %>%
  group_by(site) %>%
  summarise(mean_temp_overall = mean(mean_temp, na.rm = TRUE), .groups = "drop")
```

```{r helper-functions}
get_var <- function(model, target_group) {
  vc <- tidy(model, effects = "ran_pars", scales = "vcov")
  round(vc$estimate[vc$group == target_group], 3)
}

get_icc <- function(model) {
  vc <- as.data.frame(VarCorr(model))
  yr_var <- vc$vcov[vc$grp == "year"]
  sy_var <- vc$vcov[vc$grp == "siteyear"]

  if (length(sy_var) == 0) {
    sy_var <- vc$vcov[vc$grp == "Residual"]
  }

  val <- yr_var / (yr_var + sy_var)
  return(as.numeric(val))
}

get_year_var <- function(model) {
  vc <- as.data.frame(VarCorr(model))
  yr_var <- vc$vcov[vc$grp == "year"]
  return(as.numeric(yr_var))
}
```
***
 
# Results

# 1. Temporal variation in breeding traits

```{r annual-summaries}
annual_fed <- FED_site_medians_exc2020 %>%
  group_by(year) %>%
  summarise(grand = mean(median_fed, na.rm = TRUE),
            q25   = quantile(median_fed, 0.25, na.rm = TRUE),
            q75   = quantile(median_fed, 0.75, na.rm = TRUE))

annual_cs <- site_means %>%
  group_by(year) %>%
  summarise(grand = mean(mean_cs, na.rm = TRUE),
            q25   = quantile(mean_cs, 0.25, na.rm = TRUE),
            q75   = quantile(mean_cs, 0.75, na.rm = TRUE))

annual_suc <- site_means %>%
  group_by(year) %>%
  summarise(grand = mean(mean_suc, na.rm = TRUE),
            q25   = quantile(mean_suc, 0.25, na.rm = TRUE),
            q75   = quantile(mean_suc, 0.75, na.rm = TRUE))
```


```{r figure-3-spaghetti, fig.width = 10, fig.height = 11, fig.cap = "**Figure 3.** Temporal variation in (a) median lay date (n = 1719), (b) mean clutch size (n = 2107), and (c) mean fledgling success (n = 1833) across the Phenoweb (43 blue tit nest box sites in Scotland), 2014-2025. Grey lines show individual site trajectories. Coloured lines and points show the annual grand median (a) or mean (b, c) across all sites. Shaded ribbons show the interquartile range. Year 2020 is excluded from lay date due to reduced fieldwork."}

theme_diss <- theme_classic(base_size = 12, base_family = "Palatino") +
  theme(axis.text.x    = element_text(angle = 45, hjust = 1, size = 9),
        axis.title.y   = element_text(size = 10),
        axis.title.x   = element_text(size = 10),
        plot.margin    = margin(5, 10, 5, 10))

p_fed <- ggplot() +
  geom_line(data = FED_site_medians_exc2020,
            aes(x = year, y = median_fed, group = site),
            colour = "grey75", alpha = 0.5, linewidth = 0.35) +
  geom_ribbon(data = annual_fed,
              aes(x = year, ymin = q25, ymax = q75),
              fill = "steelblue", alpha = 0.15) +
  geom_line(data = annual_fed,
            aes(x = year, y = grand),
            colour = "steelblue", linewidth = 1.1) +
  geom_point(data = annual_fed,
             aes(x = year, y = grand),
             colour = "steelblue", size = 1.8) +
  scale_x_continuous(breaks = seq(2014, 2025, by = 1)) +
  scale_y_continuous(breaks = pretty_breaks(n = 6)) +
  labs(x = NULL, y = "Median Lay date\n(day of year)") +
  theme_diss

p_cs <- ggplot() +
  geom_line(data = site_means,
            aes(x = year, y = mean_cs, group = site),
            colour = "grey75", alpha = 0.5, linewidth = 0.35) +
  geom_ribbon(data = annual_cs,
              aes(x = year, ymin = q25, ymax = q75),
              fill = "#57A85A", alpha = 0.15) +
  geom_line(data = annual_cs,
            aes(x = year, y = grand),
            colour = "#57A85A", linewidth = 1.1) +
  geom_point(data = annual_cs,
             aes(x = year, y = grand),
             colour = "#57A85A", size = 1.8) +
  scale_x_continuous(breaks = seq(2014, 2025, by = 1)) +
  scale_y_continuous(breaks = pretty_breaks(n = 6)) +
  labs(x = NULL, y = "Mean clutch size\n(eggs)") +
  theme_diss

p_suc <- ggplot() +
  geom_line(data = site_means,
            aes(x = year, y = mean_suc, group = site),
            colour = "grey75", alpha = 0.5, linewidth = 0.35) +
  geom_ribbon(data = annual_suc,
              aes(x = year, ymin = q25, ymax = q75),
              fill = "#E07B4F", alpha = 0.15) +
  geom_line(data = annual_suc,
            aes(x = year, y = grand),
            colour = "#E07B4F", linewidth = 1.1) +
  geom_point(data = annual_suc,
             aes(x = year, y = grand),
             colour = "#E07B4F", size = 1.8) +
  scale_x_continuous(breaks = seq(2014, 2025, by = 1)) +
  scale_y_continuous(breaks = pretty_breaks(n = 6)) +
  labs(x = "Year", y = "Mean fledgling success\n(fledglings)") +
  theme_diss

p_fed / p_cs / p_suc +
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(", tag_suffix = ")") &
  theme(plot.tag = element_text(size = 12, face = "bold"))
```

***

# 2. Overall spatial synchrony — Aim 1

## 2.1 Baseline spatial synchrony

```{r pairwise-matrices}
mat_fed <- site_means %>%
  filter(year != 2020) %>%
  select(site, year, median_fed) %>%
  pivot_wider(names_from = site, values_from = median_fed) %>%
  arrange(year) %>% column_to_rownames("year") %>% as.matrix()

mat_cs <- site_means %>%
  select(site, year, mean_cs) %>%
  pivot_wider(names_from = site, values_from = mean_cs) %>%
  arrange(year) %>% column_to_rownames("year") %>% as.matrix()

mat_suc <- site_means %>%
  select(site, year, mean_suc) %>%
  pivot_wider(names_from = site, values_from = mean_suc) %>%
  arrange(year) %>% column_to_rownames("year") %>% as.matrix()
```

```{r pairwise-combined}
Pairwise_combined <- Pairwise_correlations %>%
  separate(site_pair, into = c("Site1", "Site2"), sep = "-", remove = FALSE) %>%
  filter(Site1 != "TOM", Site2 != "TOM") %>%
  rowwise() %>%
  mutate(
    n_shared_fed = sum(complete.cases(cbind(mat_fed[, Site1], mat_fed[, Site2]))),
    n_shared_cs  = sum(complete.cases(cbind(mat_cs[,  Site1], mat_cs[,  Site2]))),
    n_shared_suc = sum(complete.cases(cbind(mat_suc[, Site1], mat_suc[, Site2])))
  ) %>%
  ungroup() %>%
  select(site_pair, Site1, Site2, distance_km,
         cor_fed, n_shared_fed, cor_cs, n_shared_cs, cor_suc, n_shared_suc)
```

```{r avg-pairwise-cors}
avg_pairwise_cors <- Pairwise_combined %>%
  summarise(
    mean_cor_fed = mean(cor_fed, na.rm = TRUE),
    mean_cor_cs  = mean(cor_cs,  na.rm = TRUE),
    mean_cor_suc = mean(cor_suc, na.rm = TRUE)
  )
```

```{r fit-baseline-models}
# ── Table 1 baseline models (full per-trait data) ──

# Sample sizes for captions
n_fed_t1 <- FED_Blue_tits_exc_2020 %>% drop_na(fed) %>% nrow()
n_cs_t1  <- Blue_tits_final %>% drop_na(cs) %>% nrow()
n_suc_t1 <- Blue_tits_final %>% drop_na(suc) %>% nrow()

# 1. Create subset dataframes with NO missing values for the target trait
fed_data <- FED_Blue_tits_exc_2020 %>% drop_na(fed)
cs_data  <- Blue_tits_final %>% drop_na(cs)
suc_data <- Blue_tits_final %>% drop_na(suc)

# 2. Fit the models on these explicitly cleaned dataframes
fed_base_model <- lmer(fed ~ (1|year) + (1|site) + (1|siteyear), data = fed_data)
cs_base_model  <- lmer(cs  ~ (1|year) + (1|site) + (1|siteyear), data = cs_data)
suc_base_model <- lmer(suc ~ (1|year) + (1|site) + (1|siteyear), data = suc_data)
```

```{r icc-bootstraps, cache=TRUE}
set.seed(42)
# ── Bootstrap CIs using lmeresampler ──
# Runs once and saves to file; subsequent knits load from file.
# Delete icc_boot_results.rds to rerun (e.g. with higher B).

if (!file.exists("icc_boot_results.rds")) {
  B <- 5000

  boot_fed_icc  <- lmeresampler::bootstrap(fed_base_model, .f = get_icc,      type = "parametric", B = B)
  boot_cs_icc   <- lmeresampler::bootstrap(cs_base_model,  .f = get_icc,      type = "parametric", B = B)
  boot_suc_icc  <- lmeresampler::bootstrap(suc_base_model, .f = get_icc,      type = "parametric", B = B)

  boot_fed_year <- lmeresampler::bootstrap(fed_base_model, .f = get_year_var, type = "parametric", B = B)
  boot_cs_year  <- lmeresampler::bootstrap(cs_base_model,  .f = get_year_var, type = "parametric", B = B)
  boot_suc_year <- lmeresampler::bootstrap(suc_base_model, .f = get_year_var, type = "parametric", B = B)

  saveRDS(list(
    boot_fed_icc  = boot_fed_icc,
    boot_cs_icc   = boot_cs_icc,
    boot_suc_icc  = boot_suc_icc,
    boot_fed_year = boot_fed_year,
    boot_cs_year  = boot_cs_year,
    boot_suc_year = boot_suc_year,
    B = B
  ), "icc_boot_results.rds")
}

icc_boots <- readRDS("icc_boot_results.rds")
boot_fed_icc  <- icc_boots$boot_fed_icc
boot_cs_icc   <- icc_boots$boot_cs_icc
boot_suc_icc  <- icc_boots$boot_suc_icc
boot_fed_year <- icc_boots$boot_fed_year
boot_cs_year  <- icc_boots$boot_cs_year
boot_suc_year <- icc_boots$boot_suc_year
```

```{r ci-helpers}
# ── Extract 95% CI using the official lmeresampler method ──
get_ci_str <- function(boot_obj) {
  
  # The package's built-in confint function handles its own internal structure.
  # We use "perc" (percentile) because it's best for ICCs bounded 0-1.
  ci <- confint(boot_obj, type = "perc")
  
  # Format it into your [Lower, Upper] brackets
  paste0("[", round(ci$lower, 3), ", ", round(ci$upper, 3), "]")
}

get_year_var_ci_str <- function(boot_obj) {
  # lmeresampler stores replicates as a data frame; extract the first (only) column
  rep_df <- boot_obj$replicates
  if (is.data.frame(rep_df)) {
    vals <- as.numeric(rep_df[[1]])
  } else {
    vals <- as.numeric(rep_df)
  }
  # Drop any NA / degenerate replicates
  vals <- vals[is.finite(vals)]
  ci <- quantile(vals, probs = c(0.025, 0.975), na.rm = TRUE)
  paste0("[", round(ci[1], 3), ", ", round(ci[2], 3), "]")
}
```


```{r table-1-baseline-synchrony}
table_1 <- tibble(
  Trait = c("Lay date", "Clutch size", "Fledgling success"),
  `$\\sigma^2_{\\text{year}}$` = c(
    get_var(fed_base_model, "year"),
    get_var(cs_base_model, "year"),
    get_var(suc_base_model, "year")
  ),
  `Year var 95% CI` = c(
    get_year_var_ci_str(boot_fed_year),
    get_year_var_ci_str(boot_cs_year),
    get_year_var_ci_str(boot_suc_year)
  ),
  `$\\sigma^2_{\\text{site}}$` = c(
    get_var(fed_base_model, "site"),
    get_var(cs_base_model, "site"),
    get_var(suc_base_model, "site")
  ),
  `$\\sigma^2_{\\text{site:year}}$` = c(
    get_var(fed_base_model, "siteyear"),
    get_var(cs_base_model, "siteyear"),
    get_var(suc_base_model, "siteyear")
  ),
  `$\\sigma^2_{\\text{resid}}$` = c(
    get_var(fed_base_model, "Residual"),
    get_var(cs_base_model, "Residual"),
    get_var(suc_base_model, "Residual")
  ),
  ICC = round(c(
    get_icc(fed_base_model),
    get_icc(cs_base_model),
    get_icc(suc_base_model)
  ), 3),
  `95% CI` = c(
    get_ci_str(boot_fed_icc),
    get_ci_str(boot_cs_icc),
    get_ci_str(boot_suc_icc)
  ),
  `Mean r` = round(c(
    avg_pairwise_cors$mean_cor_fed,
    avg_pairwise_cors$mean_cor_cs,
    avg_pairwise_cors$mean_cor_suc
  ), 3)
)

kable(
  table_1,
  format = "html",
  align = "lcccccccc",
  escape = FALSE,
  caption = paste0(
    "**Table 1.** Baseline synchrony estimates and variance components ",
    "from raw nest-level LMMs (FED: n = ", n_fed_t1,
    "; CS: n = ", n_cs_t1,
    "; SUC: n = ", n_suc_t1,
    "). 95% CI from parametric bootstrap (B = 5000). ",
    "Mean r = average pairwise Pearson r from detrended site time series."
  )
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

```{r trait-relationship-data}
ICC_estimates_CS <- FED_Blue_tits_exc_2020 %>%
  select(year, site, box, fed, cs, suc, siteyear) %>%
  drop_na(fed, cs)

ICC_estimates_SUC1 <- FED_Blue_tits_exc_2020 %>%
  select(year, site, box, fed, cs, suc, siteyear) %>%
  drop_na(fed, suc)

ICC_estimates_SUC2 <- Blue_tits_final %>%
  select(year, site, box, fed, cs, suc, siteyear) %>%
  drop_na(cs, suc)

ICC_estimates_SUC3 <- FED_Blue_tits_exc_2020 %>%
  select(year, site, box, fed, cs, suc, siteyear) %>%
  drop_na(fed, cs, suc)
```

```{r trait-lms}
fed_cs_lm  <- lm(cs ~ fed,  data = ICC_estimates_CS)
fed_suc_lm <- lm(suc ~ fed, data = ICC_estimates_SUC1)
cs_suc_lm  <- lm(suc ~ cs,  data = ICC_estimates_SUC2)
```

## 2.2 Synchrony after controlling for upstream traits

As described in the methods, breeding traits form a biological cascade and are therefore sequentially dependent (Figure A2); when tested with our data, lay date had an effect on clutch size and fledgling success, and clutch size had an effect on fledgling success. Accordingly, I re-estimated synchrony values after controlling for upstream traits (adding them as fixed effects in later models).

```{r icc-raw-fixed-effect-models}
# CS controlling for FED
cs_fixed_fed_model <- lmer(cs ~ fed + (1|year) + (1|site) + (1|siteyear), ICC_estimates_CS)

# SUC controlling for FED + CS
suc_fixed_fed.cs_model <- lmer(suc ~ fed + cs + (1|year) + (1|site) + (1|siteyear), ICC_estimates_SUC3)
```

```{r icc-bootstraps-t2, cache=TRUE}
set.seed(42)
# ── Table 2 baseline models (matched subsets) ──
cs_base_model_t2  <- lmer(cs ~ (1|year) + (1|site) + (1|siteyear), data = ICC_estimates_CS)
suc_base_model_t2 <- lmer(suc ~ (1|year) + (1|site) + (1|siteyear), data = ICC_estimates_SUC3)

# Sample sizes for caption
n_cs_t2  <- nrow(ICC_estimates_CS)
n_suc_t2 <- nrow(ICC_estimates_SUC3)

# ── Bootstrap CIs for Table 2 models ──
# NOTE: Delete icc_boot_t2_results.rds before re-knitting to regenerate with
# the year variance bootstraps included.
if (!file.exists("icc_boot_t2_results.rds")) {
  B2 <- 5000
  
  # Table 2 matched-subset baselines — ICC
  boot_cs_base_t2      <- lmeresampler::bootstrap(cs_base_model_t2,       .f = get_icc,      type = "parametric", B = B2)
  boot_suc_base_t2     <- lmeresampler::bootstrap(suc_base_model_t2,      .f = get_icc,      type = "parametric", B = B2)
  
  # Table 2 with fixed effects — ICC
  boot_cs_fed_icc      <- lmeresampler::bootstrap(cs_fixed_fed_model,     .f = get_icc,      type = "parametric", B = B2)
  boot_suc_fed.cs_icc  <- lmeresampler::bootstrap(suc_fixed_fed.cs_model, .f = get_icc,      type = "parametric", B = B2)

  # Table 2 — year variance CIs
  boot_cs_base_year_t2  <- lmeresampler::bootstrap(cs_base_model_t2,       .f = get_year_var, type = "parametric", B = B2)
  boot_cs_fed_year      <- lmeresampler::bootstrap(cs_fixed_fed_model,     .f = get_year_var, type = "parametric", B = B2)
  boot_suc_base_year_t2 <- lmeresampler::bootstrap(suc_base_model_t2,      .f = get_year_var, type = "parametric", B = B2)
  boot_suc_fed.cs_year  <- lmeresampler::bootstrap(suc_fixed_fed.cs_model, .f = get_year_var, type = "parametric", B = B2)
  
  saveRDS(list(
    boot_cs_base_t2       = boot_cs_base_t2,
    boot_suc_base_t2      = boot_suc_base_t2,
    boot_cs_fed_icc       = boot_cs_fed_icc,
    boot_suc_fed.cs_icc   = boot_suc_fed.cs_icc,
    boot_cs_base_year_t2  = boot_cs_base_year_t2,
    boot_cs_fed_year      = boot_cs_fed_year,
    boot_suc_base_year_t2 = boot_suc_base_year_t2,
    boot_suc_fed.cs_year  = boot_suc_fed.cs_year,
    B2 = B2
  ), "icc_boot_t2_results.rds")
}

t2_boots <- readRDS("icc_boot_t2_results.rds")
boot_cs_base_t2       <- t2_boots$boot_cs_base_t2
boot_suc_base_t2      <- t2_boots$boot_suc_base_t2
boot_cs_fed_icc       <- t2_boots$boot_cs_fed_icc
boot_suc_fed.cs_icc   <- t2_boots$boot_suc_fed.cs_icc
boot_cs_base_year_t2  <- t2_boots$boot_cs_base_year_t2
boot_cs_fed_year      <- t2_boots$boot_cs_fed_year
boot_suc_base_year_t2 <- t2_boots$boot_suc_base_year_t2
boot_suc_fed.cs_year  <- t2_boots$boot_suc_fed.cs_year
```

```{r figure-density-plot-data-prep1}

density_data <- bind_rows(
  tibble(icc = as.numeric(boot_fed_icc$replicates),        trait = "Lay date",         model = "Baseline"),
  tibble(icc = as.numeric(boot_cs_base_t2$replicates),         trait = "Clutch size",        model = "Baseline"),
  tibble(icc = as.numeric(boot_cs_fed_icc$replicates),     trait = "Clutch size",        model = "+ FED"),
  tibble(icc = as.numeric(boot_suc_base_t2$replicates),        trait = "Fledgling success",  model = "Baseline"),
  tibble(icc = as.numeric(boot_suc_fed.cs_icc$replicates), trait = "Fledgling success",  model = "+ FED + CS")
) %>%
  mutate(
    trait = factor(trait, levels = c("Lay date", "Clutch size", "Fledgling success")),
    model = factor(model, levels = c("Baseline", "+ FED", "+ FED + CS"))
  )

point_estimates <- tibble(
  trait = factor(c("Lay date", "Clutch size", "Clutch size",
                   "Fledgling success", "Fledgling success"),
                 levels = c("Lay date", "Clutch size", "Fledgling success")),
  model = factor(c("Baseline", "Baseline", "+ FED", "Baseline", "+ FED + CS"),
                 levels = c("Baseline", "+ FED", "+ FED + CS")),
  icc   = c(get_icc(fed_base_model),
            get_icc(cs_base_model_t2),   get_icc(cs_fixed_fed_model),
            get_icc(suc_base_model_t2),  get_icc(suc_fixed_fed.cs_model))
)

trait_cols <- c("Lay date"         = "steelblue",
                "Clutch size"       = "#57A85A",
                "Fledgling success" = "#E07B4F")
```

```{r figure-density-plot-data-prep2}
plot_density_data <- density_data %>%
  filter(!(grepl("Fledgling", trait, ignore.case = TRUE) & model == "+ FED"))

plot_point_est <- point_estimates %>%
  filter(!(grepl("Fledgling", trait, ignore.case = TRUE) & model == "+ FED"))


density_baseline <- plot_density_data %>% filter(model == "Baseline")
density_fixed    <- plot_density_data %>% filter(model != "Baseline")

```

```{r figure-density-plot, fig.width = 6, fig.height = 6, fig.cap = paste0("**Figure 4.** Bootstrap density distributions of ICC estimates from parametric bootstrapping (B =  5000) using matching subsets of data. (a) Lay date (n = 1719)— baseline only. (b) Clutch size (n = 1700) — baseline (solid) vs controlling for lay date (dashed). (c) Fledgling success (n = 1391) — baseline (solid) vs controlling for lay date and clutch size (dot-dash).")}

ggplot() +
  geom_density(data = density_baseline %>% filter(icc >= 0 & icc <= 1),
               aes(x = icc, colour = trait, group = trait, linetype = "Baseline", fill = after_scale(alpha(colour, 0.15))),
               alpha = 0.15, linewidth = 0.5, show.legend = TRUE) +
  
  geom_density(data = density_fixed %>% filter(icc >= 0 & icc <= 1),
               aes(x = icc, colour = trait, linetype = model, group = interaction(trait, model)),
               linewidth = 0.5, fill = NA, key_glyph = "path") +

  geom_vline(data = plot_point_est,
             aes(xintercept = icc, colour = trait, linetype = model),
             linewidth = 0.3, show.legend = FALSE) +
  
geom_text(data = plot_point_est %>% filter(model == "Baseline"),
          aes(x = icc, y = Inf, label = paste0("ICC = ", sprintf("%.3f", icc))),
          colour = "black",
          vjust = 1.5, hjust = -0.1, size = 3, fontface = "bold",
          show.legend = FALSE) +
geom_text(data = plot_point_est %>% filter(model != "Baseline"),
          aes(x = icc, y = Inf, label = paste0("ICC = ", sprintf("%.3f", icc))),
          colour = "grey40",
          vjust = 3.5, hjust = -0.1, size = 3, fontface = "italic",
          show.legend = FALSE) +
  
  facet_wrap(~ trait, ncol = 1) +
  scale_colour_manual(values = trait_cols, guide = "none") +
  scale_fill_manual(values = trait_cols, guide = "none") +
  
 scale_linetype_manual(
    values = c("Baseline" = "solid", "+ FED" = "dashed", "+ FED + CS" = "dashed"),
    breaks = c("Baseline", "+ FED"),
    labels = c("Baseline", "+ Upstream traits"),
    name = "Model:"
  ) +
guides(linetype = guide_legend(
    keywidth = unit(0.8, "cm"),
    keyheight = unit(0.3, "cm"),
    override.aes = list(
      fill    = NA,
      alpha   = 1,
      linewidth = 0.5,
      colour  = "black"
    )
  )) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.25)) +
  labs(x = "ICC (synchrony estimate)", y = "Density") +
  theme_classic(base_size = 8, base_family = "Palatino") +
  theme(
    strip.background = element_rect(fill = "grey92", colour = NA),
    strip.text       = element_text(face = "bold", size = 8),
    legend.position  = "bottom"
  )
```
***

```{r table-2-fixed-effects-with-variance}
table_2 <- tibble(
  Trait = c("Clutch size", "", "Fledgling success", ""),
  `Fixed effect` = c(
    "None", "Laying date",
    "None", "Laying date + Clutch size"),
  `$\\sigma^2_{\\text{year}}$` = c(
    get_var(cs_base_model_t2, "year"),
    get_var(cs_fixed_fed_model, "year"),
    get_var(suc_base_model_t2, "year"),
    get_var(suc_fixed_fed.cs_model, "year")),
  `Year var 95% CI` = c(
    get_year_var_ci_str(boot_cs_base_year_t2),
    get_year_var_ci_str(boot_cs_fed_year),
    get_year_var_ci_str(boot_suc_base_year_t2),
    get_year_var_ci_str(boot_suc_fed.cs_year)
  ),
  `$\\sigma^2_{\\text{site}}$` = c(
    get_var(cs_base_model_t2, "site"),
    get_var(cs_fixed_fed_model, "site"),
    get_var(suc_base_model_t2, "site"),
    get_var(suc_fixed_fed.cs_model, "site")),
  `$\\sigma^2_{\\text{site:year}}$` = c(
    get_var(cs_base_model_t2, "siteyear"),
    get_var(cs_fixed_fed_model, "siteyear"),
    get_var(suc_base_model_t2, "siteyear"),
    get_var(suc_fixed_fed.cs_model, "siteyear")),
  `$\\sigma^2_{\\text{resid}}$` = c(
    get_var(cs_base_model_t2, "Residual"),
    get_var(cs_fixed_fed_model, "Residual"),
    get_var(suc_base_model_t2, "Residual"),
    get_var(suc_fixed_fed.cs_model, "Residual")),
  `ICC` = round(c(
    get_icc(cs_base_model_t2),
    get_icc(cs_fixed_fed_model),
    get_icc(suc_base_model_t2),
    get_icc(suc_fixed_fed.cs_model)), 3),
  `95% CI` = c(
    get_ci_str(boot_cs_base_t2),
    get_ci_str(boot_cs_fed_icc),
    get_ci_str(boot_suc_base_t2),
    get_ci_str(boot_suc_fed.cs_icc))
)

kable(table_2, format = "html", align = "llccccccc", escape = FALSE,
      caption = paste0("**Table 2.** Effect of controlling for upstream traits on variance ",
                       "components and ICC. Baseline models refitted on matched subsets ",
                       "(Clutch size: n = ", n_cs_t2, "; Fledgling success: n = ", n_suc_t2, 
                       ") for direct comparability with fixed-effect models. ",
                       "Full-sample baseline estimates in Table 1. ",
                       "95% CI from parametric bootstrap (B = ", t2_boots$B2, "). ")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  pack_rows("Clutch size", 1, 2) %>%
  pack_rows("Fledgling success", 3, 4)
```


