---
title: "07 — Weather drivers of synchrony"
author: "Alexander Hobbs"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: flatly
---

<!--
Tests whether mean spring temperature and mean spring precipitation
(Feb-May, site-specific from HadUK-Grid; Hollis et al. 2019) explain the
observed inter-annual synchrony in lay date. Adds each weather variable
as a fixed effect within the same LMM structure as 04_results_synchrony_iccs.Rmd
and re-estimates the year, site, site:year and residual variance components.
Reproduces Table 4 in the dissertation.

Inputs (data/, produced by 01_data_cleaning.Rmd and earlier scripts):
  - Blue_tits_cleaned.csv
  - site_means.csv, site_details.csv
  - Pairwise_correlations.csv (precomputed pairwise Pearson r)
  - icc_baseline.csv (output from 04_results_synchrony_iccs.Rmd)
  - site_means_with_temp.csv, climate_data.csv
  - daily_maxtemp_Hollis.txt
  - daily_mintemp_Hollis.txt
  - daily_precipitation_Hollis.txt

Bootstrapping: parametric bootstraps (B3 = 5000) cached as
  icc_boot_t3_results.rds. Random seed 42 set before the bootstrap chunk
  for reproducibility.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      fig.align = "center", dpi = 300)
```

```{r load-libraries}
library(tidyverse)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(broom)
library(knitr)
library(kableExtra)
library(corrr)
library(reshape2)
library(ncf)
library(mvtnorm)
library(matrixStats)
library(progress)
library(patchwork)
library(scales)
library(ggforce)
library(viridis)
library(geosphere)
```

***

# Data loading and cleaning

```{r load-data}
site_means <- read.csv(
  "data/site_means.csv",
  stringsAsFactors = FALSE, header = TRUE)

site_details <- read.csv(
  "data/site_details.csv",
  stringsAsFactors = FALSE, header = TRUE)

Blue_tits_cleaned <- read.csv(
  'data/Blue_tits_cleaned.csv',
  stringsAsFactors = TRUE, header = TRUE)

Pairwise_correlations <- read.csv(
  "data/Pairwise_correlations.csv",
  stringsAsFactors = FALSE, header = TRUE)

icc_baseline <- read.csv(
  "data/icc_baseline.csv")

site_means_with_temp <- read.csv(
  "data/site_means_with_temp.csv",
  stringsAsFactors = FALSE, header = TRUE)

climate_data <- read.csv(
  "data/climate_data.csv",
  stringsAsFactors = FALSE, header = TRUE)

max_temp <- read.table("data/daily_maxtemp_Hollis.txt", header = TRUE)

min_temp <- read.table("data/daily_mintemp_Hollis.txt", header = TRUE)

# Daily precipitation — wide format, columns "1":"366" = day of year
precip_raw <- read.table(
  "data/daily_precipitation_Hollis.txt",
  header = TRUE, stringsAsFactors = FALSE, check.names = FALSE)
```


```{r data-cleaning}
site_means <- site_means %>% filter(site != "TOM")
site_details <- site_details %>% filter(site != "TOM")

Blue_tits_final <- Blue_tits_cleaned %>% filter(site != "TOM")
Blue_tits_final$siteyear <- paste(Blue_tits_final$site, Blue_tits_final$year)

FED_site_medians_exc2020 <- site_means %>% filter(year != 2020)
FED_Blue_tits_exc_2020 <- Blue_tits_final %>% filter(year != 2020)
FED_Blue_tits_exc_2020$siteyear <- paste(FED_Blue_tits_exc_2020$site, 
                                          FED_Blue_tits_exc_2020$year)

site_details_clean <- site_details %>%
  dplyr::select(site, Mean.Lat, Mean.Long) %>%
  dplyr::rename(Lat = Mean.Lat, Lon = Mean.Long)

Blue_tits_final <- Blue_tits_final %>%
  left_join(climate_data %>% select(site, year, mean_temp), by = c("site", "year")) %>%
  mutate(mean_temp_z = scale(mean_temp)[, 1])

FED_Blue_tits_exc_2020 <- FED_Blue_tits_exc_2020 %>%
  left_join(climate_data %>% select(site, year, mean_temp), by = c("site", "year")) %>%
  mutate(mean_temp_z = scale(mean_temp)[, 1])

site_temp_summary <- climate_data %>%
  group_by(site) %>%
  summarise(mean_temp_overall = mean(mean_temp, na.rm = TRUE), .groups = "drop")

feb_may_cols <- as.character(32:151)

precip_spring <- precip_raw %>%
  filter(site != "TOM") %>%
  mutate(mean_precip = rowMeans(select(., all_of(feb_may_cols)), na.rm = TRUE)) %>%
  select(site, year, mean_precip)

FED_Blue_tits_exc_2020 <- FED_Blue_tits_exc_2020 %>%
  left_join(precip_spring, by = c("site", "year")) %>%
  mutate(mean_precip_z = scale(mean_precip)[, 1])

```


# 4. Testing the effects of climate on synchrony - Aim 3

Distance-decay is not present within the study's extent (figure 5), indicating that the primary driver of synchrony within breeding traits operates regionally or over the study transect. To test whether mean spring temperature and precipitation are contributing to this observed pattern, I added each climate variable as a fixed effect within separate LMMs for lay date.

## 4.1  climate as fixed effects in ICC models

```{r helper-functions}
# ── Variance extraction ──
get_var <- function(model, group) {
  vc <- as.data.frame(VarCorr(model))
  vc$vcov[vc$grp == group]
}

# ── ICC calculation ──
get_icc <- function(model) {
  vc <- as.data.frame(VarCorr(model))
  yr_var <- vc$vcov[vc$grp == "year"]
  sy_var <- vc$vcov[vc$grp == "siteyear"]
  if (length(sy_var) == 0) {
    sy_var <- vc$vcov[vc$grp == "Residual"]
  }
  val <- yr_var / (yr_var + sy_var)
  return(as.numeric(val))
}

# ── CI extraction from lmeresampler bootstrap ──
get_ci_str <- function(boot_obj) {
  # Extract the bootstrapped ICC values
  boot_vals <- unlist(boot_obj$replicates)
  ci <- quantile(boot_vals, probs = c(0.025, 0.975), na.rm = TRUE)
  paste0("[", round(ci[1], 3), ", ", round(ci[2], 3), "]")
}

get_year_var <- function(model) {
  vc <- as.data.frame(VarCorr(model))
  yr_var <- vc$vcov[vc$grp == "year"]
  return(as.numeric(yr_var))
}

get_year_var_ci_str <- function(boot_obj) {
  rep_df <- boot_obj$replicates
  if (is.data.frame(rep_df)) {
    vals <- as.numeric(rep_df[[1]])
  } else {
    vals <- as.numeric(rep_df)
  }
  vals <- vals[is.finite(vals)]
  ci <- quantile(vals, probs = c(0.025, 0.975), na.rm = TRUE)
  paste0("[", round(ci[1], 3), ", ", round(ci[2], 3), "]")
}

```

```{r aim3-matched-subset}
# Matched subset — rows with fed, temperature, and precipitation all present
# so all three models (baseline, temp, precip) are fitted on the same data
fed_temp_data <- FED_Blue_tits_exc_2020 %>% drop_na(fed, mean_temp_z, mean_precip_z)

# Baseline on matched subset
fed_base_t3 <- lmer(fed ~ (1|year) + (1|site) + (1|siteyear), data = fed_temp_data)

# Temperature model
fed_temp_t3 <- lmer(fed ~ mean_temp_z + (1|year) + (1|site) + (1|siteyear), data = fed_temp_data)

# Precipitation model
fed_precip_t3 <- lmer(fed ~ mean_precip_z + (1|year) + (1|site) + (1|siteyear), data = fed_temp_data)

n_fed_t3 <- nrow(fed_temp_data)
```
```{r}
summary(fed_temp_t3)
```


```{r aim3-bootstraps, cache=TRUE}
set.seed(42)
if (!file.exists("icc_boot_t3_results.rds")) {
  B3 <- 5000
  
  boot_fed_base_t3   <- lmeresampler::bootstrap(fed_base_t3,   .f = get_icc,      type = "parametric", B = B3)
  boot_fed_temp_t3   <- lmeresampler::bootstrap(fed_temp_t3,   .f = get_icc,      type = "parametric", B = B3)
  boot_fed_precip_t3 <- lmeresampler::bootstrap(fed_precip_t3, .f = get_icc,      type = "parametric", B = B3)

  # NEW — year variance CIs
  boot_fed_base_year_t3   <- lmeresampler::bootstrap(fed_base_t3,   .f = get_year_var, type = "parametric", B = B3)
  boot_fed_temp_year_t3   <- lmeresampler::bootstrap(fed_temp_t3,   .f = get_year_var, type = "parametric", B = B3)
  boot_fed_precip_year_t3 <- lmeresampler::bootstrap(fed_precip_t3, .f = get_year_var, type = "parametric", B = B3)
  
  saveRDS(list(
    boot_fed_base_t3        = boot_fed_base_t3,
    boot_fed_temp_t3        = boot_fed_temp_t3,
    boot_fed_precip_t3      = boot_fed_precip_t3,
    boot_fed_base_year_t3   = boot_fed_base_year_t3,
    boot_fed_temp_year_t3   = boot_fed_temp_year_t3,
    boot_fed_precip_year_t3 = boot_fed_precip_year_t3,
    B3 = B3
  ), "icc_boot_t3_results.rds")
}

t3_boots <- readRDS("icc_boot_t3_results.rds")
boot_fed_base_t3        <- t3_boots$boot_fed_base_t3
boot_fed_temp_t3        <- t3_boots$boot_fed_temp_t3
boot_fed_precip_t3      <- t3_boots$boot_fed_precip_t3
# NEW
boot_fed_base_year_t3   <- t3_boots$boot_fed_base_year_t3
boot_fed_temp_year_t3   <- t3_boots$boot_fed_temp_year_t3
boot_fed_precip_year_t3 <- t3_boots$boot_fed_precip_year_t3
```

```{r aim3-table}
temp_coef   <- summary(fed_temp_t3)$coefficients["mean_temp_z", ]
precip_coef <- summary(fed_precip_t3)$coefficients["mean_precip_z", ]

table_3 <- tibble(
  Trait = c("Lay date", "", ""),
  `Fixed effect` = c("None", "Temperature", "Precipitation"),
  `$\\sigma^2_{\\text{year}}$` = round(c(
    get_var(fed_base_t3,   "year"),
    get_var(fed_temp_t3,   "year"),
    get_var(fed_precip_t3, "year")), 3),
  `Year var 95% CI` = c(
    get_year_var_ci_str(boot_fed_base_year_t3),
    get_year_var_ci_str(boot_fed_temp_year_t3),
    get_year_var_ci_str(boot_fed_precip_year_t3)
  ),
  `$\\sigma^2_{\\text{site}}$` = round(c(
    get_var(fed_base_t3,   "site"),
    get_var(fed_temp_t3,   "site"),
    get_var(fed_precip_t3, "site")), 3),
  `$\\sigma^2_{\\text{site:year}}$` = round(c(
    get_var(fed_base_t3,   "siteyear"),
    get_var(fed_temp_t3,   "siteyear"),
    get_var(fed_precip_t3, "siteyear")), 3),
  `$\\sigma^2_{\\text{resid}}$` = round(c(
    get_var(fed_base_t3,   "Residual"),
    get_var(fed_temp_t3,   "Residual"),
    get_var(fed_precip_t3, "Residual")), 3),
  `ICC` = round(c(
    get_icc(fed_base_t3),
    get_icc(fed_temp_t3),
    get_icc(fed_precip_t3)), 3),
  `95% CI` = c(
    get_ci_str(boot_fed_base_t3),
    get_ci_str(boot_fed_temp_t3),
    get_ci_str(boot_fed_precip_t3))
)

kable(table_3, align = "llcccccc", escape = FALSE,
      caption = paste0(
        "**Table 3.** Effect of mean spring temperature and mean spring precipitation ",
        "(both Feb-May, standardised) on lay date variance ",
        "components and synchrony. All models fitted on the same ",
        "matched subset (n = ", n_fed_t3,
        ") for direct comparability. ",
        "95% CI from parametric bootstrap (B = ", t3_boots$B3, "). ",
        "2020 excluded.")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


