---
title: "Check — 2020 lay-date exclusion"
author: "Alexander Hobbs"
output: html_document
---

<!--
Diagnostic / sensitivity analysis supporting the exclusion of 2020 lay-date
data from main analyses (described in Methods 2.2).

Reduced fieldwork during COVID-19 disruption meant only ~20 first-egg-date
observations across few sites were available in 2020. This script:
  1. Quantifies the 2020 lay-date data gap.
  2. Refits the lay-date -> clutch-size cascade model with vs without 2020
     to show the practical impact on parameter estimates is negligible.

Inputs (from data/, produced by 01_data_cleaning.Rmd):
  - Blue_tits_final.csv
  - Blue_tits_cleaned.csv
  - FED_Blue_tits_exc_2020.csv
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r libraries}
library(tidyverse)
library(lme4)
library(lmerTest)
library(broom.mixed)
```

```{r load-data}
Blue_tits_final        <- read.csv("data/Blue_tits_final.csv",
                                   stringsAsFactors = FALSE)
Blue_tits_cleaned      <- read.csv("data/Blue_tits_cleaned.csv",
                                   stringsAsFactors = FALSE)
FED_Blue_tits_exc_2020 <- read.csv("data/FED_Blue_tits_exc_2020.csv",
                                   stringsAsFactors = FALSE)

# siteyear column needed by mixed models below
Blue_tits_final$siteyear        <- paste(Blue_tits_final$site,
                                          Blue_tits_final$year)
FED_Blue_tits_exc_2020$siteyear <- paste(FED_Blue_tits_exc_2020$site,
                                          FED_Blue_tits_exc_2020$year)

# site-level aggregates for site_means reference below
site_means <- Blue_tits_final %>%
  group_by(site, year) %>%
  summarise(mean_cs  = mean(cs,  na.rm = TRUE),
            mean_suc = mean(suc, na.rm = TRUE),
            .groups = "drop")
```

## 1. How much lay-date data is in 2020?

```{r investigate-2020-fed}
# Number of FED (first egg date) observations in 2020
Blue_tits_final %>%
  filter(year == 2020) %>%
  summarise(n_fed_obs   = sum(!is.na(fed)),
            n_sites     = n_distinct(site[!is.na(fed)]),
            total_sites = n_distinct(site))

# Which sites have FED data in 2020?
Blue_tits_final %>%
  filter(year == 2020, !is.na(fed)) %>%
  group_by(site) %>%
  summarise(n_obs = n(), .groups = "drop")

# What proportion of the total FED dataset is 2020?
Blue_tits_final %>%
  summarise(total_fed = sum(!is.na(fed)),
            fed_2020  = sum(!is.na(fed) & year == 2020),
            pct_2020  = round(100 * sum(!is.na(fed) & year == 2020)
                              / sum(!is.na(fed)), 1))
```

## 2. Sensitivity check: model with vs without 2020

```{r sensitivity-check-2020-fed}
# Cascade model (lay date predicting clutch size) WITH 2020 included
cs_fed_with2020 <- lmer(cs ~ fed + (1|year) + (1|site) + (1|siteyear),
                        data = Blue_tits_final %>% drop_na(fed, cs))

# Same model WITHOUT 2020
cs_fed_without2020 <- lmer(cs ~ fed + (1|year) + (1|site) + (1|siteyear),
                           data = FED_Blue_tits_exc_2020 %>% drop_na(fed, cs))

# Helper: extract ICC from the year random effect
extract_icc <- function(model) {
  v <- tidy(model, effects = "ran_pars", scales = "vcov")
  yr <- v$estimate[v$group == "year"]
  sy <- v$estimate[v$group == "siteyear"]
  round(yr / (yr + sy), 3)
}

cat("ICC with 2020:        ", extract_icc(cs_fed_with2020),   "\n")
cat("ICC without 2020:     ", extract_icc(cs_fed_without2020), "\n")
cat("FED slope with 2020:    ", round(fixef(cs_fed_with2020)["fed"],    4), "\n")
cat("FED slope without 2020: ", round(fixef(cs_fed_without2020)["fed"], 4), "\n")
```

Sensitivity analyses confirmed that inclusion or exclusion of 2020 lay-date
observations had negligible impact on model estimates, supporting the
decision to exclude 2020 lay-date data on the grounds that the small number
of observations across few sites yields unreliable site-mean estimates for
that year.

## 3. Site count sanity check

```{r site-counts}
cat("Sites before any exclusions:  ", n_distinct(Blue_tits_cleaned$site), "\n")
cat("Sites after removing TOM:     ", n_distinct(Blue_tits_final$site),  "\n")
cat("Sites in site_means:          ", n_distinct(site_means$site), "\n\n")

cat("Sites in cleaned set:\n")
sort(unique(Blue_tits_cleaned$site))
cat("\nSites in final set:\n")
sort(unique(Blue_tits_final$site))
```
